CALVIS: Chest, wAist and peLVIS circumference from 3D human body meshes as ground truth for deep learning

### 2019-10-08 PaperRMFI2019/samplepaper.tex
minor fixes

@@

\subsubsection{Axilla recognition}\label{subsec:armpit_recog}
As mentioned above, we would like to measure the chest circumference below the 
armpit, also known as axilla. This raises the question whether we can recognize 
the axilla. What we want is a point $p_a$ in the mesh at the proper location 
under 
the arms. One point suffices because we require only a reference $p^y_a$ along 
the 
y-axis to define the chest region according to equation \ref{eq:cr}. Therefore, 
we focus on the left shoulder joint $j_{ls} \in \mathbf{J}$. Since the axilla 
is related to 
the shoulder joint, we can cast a ray $\boldsymbol{\iota}$ from it in direction 
to the middle left 
inferior edge of the bounding box outside the mesh at point

\begin{align}
p_b = \bigg(v^x_{min}, v^y_{min}, v^z_{min} + 
\frac{v^z_{max} - v^z_{min}}{2}\bigg).
\end{align}

Here $v^x_{min}, v^y_{min}, v^z_{min}$ are vertex minimum $x$,$y$ and $z$ 
coordinates, respectively; analog is $v^z_{max}$ the maximum vertex $z$ 
coordinate. The ray 
$\iota$ is then determined by points $j_{ls}$ (inside the mesh) and $p_b$ 
(outside the mesh). Therefore, the ray $\iota$ intersects the mesh at one point 
$p_a$ (Figure \ref{fig:axilla-recognition}), which we consider the axilla 
center.

@@

Starting 
from the 
top $q_t$ of the bounding box we slice iteratively mesh $\mathcal{M}$ with 
plane 
$\boldsymbol{\pi_j}$ every $m$-meters along the y-axis until we reach the 
bounding box bottom 
$q_b$. Figure \ref{fig:subjects_planes_xrays} shows 40 of these 
slides. Next, we assemble the mesh sliding vector $\vec{\mathcal{L}}$. The 
elements of 
this vector are slice point y-coordinates

@@

\begin{figure}[h!]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}. See section 
		\ref{subsec:mesh_signature_extrema} for discussion of these extrema.}
	\label{fig:qualitative_eval}
\end{figure}

\subsection{Learning convergence evaluation}\label{subsec:learn_conv}
This experiment demonstrates the plausibility of using calvis to obtain data 
that can be used as a ground truth for machine learning (in general) and 
specifically as supervising signal for ANN.

Figure \ref{fig:calvis_net} shows the our prototypical ANN calvis net where the 
input is a synthetic image from a 3D mesh and the output are the three body 
dimensions.

Figure \ref{fig:learn_conv} shows that after epoch number 5 the estimation 
error decreases and the learning converges.

\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}. See section 
		\ref{subsec:mesh_signature_extrema} for discussion of these extrema.}
	\label{fig:qualitative_eval}
\end{figure}

\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}. See section 
		\ref{subsec:mesh_signature_extrema} for discussion of these extrema.}
	\label{fig:qualitative_eval}
\end{figure}

@@

The method can be optimized. Further research must be conducted.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{egbib}
%

>>>>

In general, previous work has concentrated on human body shapes not deviating 
much from the mean. In contrast, we move beyond by exploring 3D meshes 
reflecting human figures characteristics such as bulky, slim, small and 
tall subjects.

@@

\subsubsection{Axilla recognition}\label{subsec:armpit_recog}
As mentioned above, we would like to measure the chest circumference below the 
armpit, also known as axilla. This raises the question whether we can recognize 
the axilla. What we want is a point $p_a$ in the mesh at the proper location 
under 
the arms. One point suffices because we require only a reference $p^y_a$ along 
the 
y-axis to define the chest region according to equation \ref{eq:cr}. Therefore, 
we focus on the right shoulder joint $j_{rs} \in \mathbf{J}$. Since the axilla 
is related to 
the shoulder joint, we can cast a ray $\boldsymbol{\iota}$ from it in direction 
to the middle left 
inferior edge of the bounding box outside the mesh at point

\begin{align}
p_c = \bigg(v^x_{min}, v^y_{min}, v^z_{min} + 
\frac{v^z_{max} - v^z_{min}}{2}\bigg).
\end{align}

Here $v^x_{min}, v^y_{min}, v^z_{min}$ are vertex minimum $x$,$y$ and $z$ 
coordinates, respectively; analog is $v^z_{max}$ the maximum vertex $z$ 
coordinate. The ray 
$\iota$ is then determined by points $j_{rs}$ (inside the mesh) and $p_b$ 
(outside the mesh). Therefore, the ray $\iota$ intersects the mesh at one point 
$p_c$ (Figure \ref{fig:axilla-recognition}), which we consider the axilla 
center. Furthermore, we can consider an strategy to increase the method 
robustness against the large variability of the human body. Since our method 
relies on one point to establish the chest upper bound, we identify the 10 
nearest neighbors in the set of vertices $\mathcal{V}$ to $p_c$ and define 
$p_a$ to be the 
vertex with the smallest 
y coordinate of the neighbors.

@@

Starting 
from the 
top $q_t$ of the bounding box we slice iteratively mesh $\mathcal{M}$ with 
plane 
$\boldsymbol{\pi_j}$ every $m$-meters along the y-axis until we reach the 
bounding box bottom 
$q_b$. Figure \ref{fig:subjects_planes_xrays} shows 40 of these 
slides. Next, we assemble the mesh sliding vector $\vec{\mathcal{L}}$. The 
elements of 
this vector are slice points

@@

\subsection{Calvis-Net and Learning Convergence 
Evaluation}\label{subsec:learn_conv}
This experiment demonstrates the plausibility of using calvis to obtain data 
that can be used as a ground truth for machine learning (in general) and 
specifically as supervising signal for ANN.

Figure \ref{fig:calvis_net} shows our prototypical CNN Calvis-Net where the 
input is a synthetic image from a 3D mesh and the output are three body 
dimensions.

Figure \ref{fig:learning_curve} shows that after epoch number 5 the estimation 
error decreases and the learning converges.
\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Calvis-Net.png}
	\end{center}
	\caption{Calvis-Net.}
	\label{fig:calvis_net}
\end{figure}

\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Training_val_loss.png}
	\end{center}
	\caption{Learning convergence evaluation.}
	\label{fig:learning_curve}
\end{figure}

@@

The method operates optimally.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{egbib}
%
